---
title: "Rolling Stone Top 500 Albums Lists"
subtitle: "How gender diversity in the voting pool affects results"
author: "Chiara Iantorno"
institute: "Houghton University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader: 
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")

knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(error = FALSE)
```


```{r load-data, include=FALSE}
rolling_stone <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-05-07/rolling_stone.csv')
```

```{r, include=FALSE}
style_xaringan(
  title_slide_background_image = "img/confetti.jpg"
)
```


class: center, middle

## The Rolling Stone magazine has put out three top 500 albums lists. Each list has different album placements.  The 2020 list involved more female voters than the 2003 list.  

---

class: center, middle

##How did the addition of more female voters affect the ranking of the 2020 list compared to that of the 2003 list?  Did this shift in the gender of voters also reflect a change in genres between the lists?

---

class: inverse, center, middle

# List Visualizations

---

```{r, echo=FALSE}
rolling_stone_2003 <- rolling_stone %>%
  filter(!is.na(rank_2003))
```

```{r, echo=FALSE}
rolling_stone_2020 <- rolling_stone %>%
  filter(!is.na(rank_2020))
```

```{r, echo=FALSE}
gender_counts <- rolling_stone_2003 %>%
  count(artist_gender)
```

```{r, echo=FALSE}
ggplot(gender_counts, aes(x = artist_gender, y = n, fill = artist_gender)) +
  geom_bar(stat = "identity") +
  labs(title = "Gender Distribution of Artists (2003 List)",
       x = "Artist Gender",
       y = "Number of Albums") +
  theme_minimal()

```

As seen in this visualization, the vast majority of artists whose albums appear on the 2003 list are male.

---
```{r, echo=FALSE}
gender_counts2 <- rolling_stone_2020 %>%
  count(artist_gender)
```

```{r, echo=FALSE}
ggplot(gender_counts2, aes(x = artist_gender, y = n, fill = artist_gender)) +
  geom_bar(stat = "identity") +
  labs(title = "Gender Distribution of Artists (2020 List)",
       x = "Artist Gender",
       y = "Number of Albums") +
  theme_minimal()

```

Here, while there are still mostly male artists on the list, the amount of female artists with albums on the list over doubled in amount.
---

class: center, middle

---
class: inverse, middle, center

# Codes for Gender Make-up of Each List

---


---
class: inverse, middle, center

# Using xaringan

---

# xaringan

- The presentation is created using the `xaringan` package

- Use `---` to separate slides and `--` for incremental builds

--

- Like this

---

# Layouts

You can use plain text

- or bullet points

.pull-left[
or text in two columns $^*$
]
.pull-right[
- like
- this
]

.footnote[
[*] And add footnotes
]

---

# Code

```{r boring-regression}
# a boring regression
model <- lm(dist ~ speed, data = cars)
tidy(model)
glance(model)
```

---

# Plots

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
iris_modified <- iris %>%
  mutate(Species = fct_other(Species, keep = "setosa"))
```

```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() + 
  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

---

## Plot and text

.pull-left[
- Some text
- goes here
]
.pull-right[
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
]

---

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
kable(head(iris), format = "html")
```

---

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

---

background-image: url(https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80)

---

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs, e.g. &#36;\alpha+\beta$ renders $\alpha+\beta$. You can use the display style with double dollar signs:

```
$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$
```

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

Limitations:

1. The source code of a LaTeX math expression must be in one line, unless it is inside a pair of double dollar signs, in which case the starting `$$` must appear in the very beginning of a line, followed immediately by a non-space character, and the ending `$$` must be at the end of a line, led by a non-space character;

1. There should not be spaces after the opening `$` or before the closing `$`.

1. Math does not work on the title slide (see [#61](https://github.com/yihui/xaringan/issues/61) for a workaround).

---

# Feeling adventurous?

- Want to find out more about `xaringan`? See https://slides.yihui.name/xaringan/#1.

- You are welcomed to use the default styling of the slides. In fact, that's what I expect majority of you will do. You will differentiate yourself with the content of your presentation.

- But some of you might want to play around with slide styling. The 
`xaringanthemer` provides some solutions for this that: https://pkg.garrickadenbuie.com/xaringanthemer.

- And if you want more bells and whistles, there is also `xaringanExtra`: https://pkg.garrickadenbuie.com/xaringanExtra.
