---
title: "Rolling Stone Top 500 Albums Lists"
subtitle: "How gender and ethnic/racial diversity in the voting pool affects results"
author: "Chiara Iantorno"
institute: "Houghton University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader: 
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")

knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(error = FALSE)
```


```{r load-data, include=FALSE}
rolling_stone <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-05-07/rolling_stone.csv')
```

```{r, include=FALSE}
style_xaringan(
  title_slide_background_image = "https://images.unsplash.com/photo-1588532218970-c2cab983746a?q=80&w=3087&auto=format&fit=crop&ixlib=rb-4.0.3&ixid=M3wxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8fA%3D%3D"
)
```


class: center, middle

## The Rolling Stone magazine has put out three top 500 albums lists. Each list has different album placements.  The 2020 list involved more female voters than the 2003 list.  It also involved more voters of color.

---

class: center, middle

##How did the addition of more female voters affect the ranking of the 2020 list compared to that of the 2003 list?  Did this shift in the gender of voters also reflect a change in genres between the lists?

---

class: inverse, center, middle

# Gender List Visualizations

---

```{r, echo=FALSE}
rolling_stone_2003 <- rolling_stone %>%
  filter(!is.na(rank_2003))
```

```{r, echo=FALSE}
rolling_stone_2020 <- rolling_stone %>%
  filter(!is.na(rank_2020))
```

```{r, echo=FALSE}
gender_counts_2003 <- rolling_stone_2003 %>%
  count(artist_gender)
```

```{r, echo=FALSE}
ggplot(gender_counts_2003, aes(x = artist_gender, y = n, fill = artist_gender)) +
  geom_bar(stat = "identity") +
  labs(title = "Gender Distribution of Artists (2003 List)",
       x = "Artist Gender",
       y = "Number of Albums") +
  theme_minimal()

```

- Most artists are male
- Not many female artists appear

---
```{r, echo=FALSE}
gender_counts_2020 <- rolling_stone_2020 %>%
  count(artist_gender)
```

```{r, echo=FALSE}
ggplot(gender_counts_2020, aes(x = artist_gender, y = n, fill = artist_gender)) +
  geom_bar(stat = "identity") +
  labs(title = "Gender Distribution of Artists (2020 List)",
       x = "Artist Gender",
       y = "Number of Albums") +
  theme_minimal()

```

- Majority of artists are still male
- Female artists have over doubled
---

class: inverse, middle, center

# Tables for Gender Make-up of Each List

---


.pull-left[
```{r, echo = FALSE}
kable(head(gender_counts_2003), format = "html")
```
]
.pull-right[
```{r, echo = FALSE}
kable(head(gender_counts_2020), format = "html")
```
]

.pull-left[
# 2003 list
]

.pull-right[
# 2020 list
]


---

.pull-left[
```{r, echo=FALSE}
rolling_stone_2003 %>%
  count(artist_gender) %>%
  mutate(proportion = percent(n / sum(n))) %>%
  kable(format = "html", col.names = c("Artist Gender", "Album Count", "Proportion"))
```
]

.pull-right[
```{r, echo=FALSE}
rolling_stone_2020 %>%
  count(artist_gender) %>%
  mutate(proportion = percent(n / sum(n))) %>%
  kable(format = "html", col.names = c("Artist Gender", "Album Count", "Proportion"))
```
]


.pull-left[
# 2003 list
]

.pull-right[
# 2020 list
]
---

class: inverse, middle, center

# Genre List Visualizations

---

class: center, middle

#As mentioned earlier, the pool of voters for the 2020 list was more racially diverse.  Some music genres, like Hip-hop, Latin and R&B, have roots in Black and Latin culture.  Meanwhile, genres like Rock, New Wave, and Country are more often associated with and made by White artists.

---

class: center middle

#Could the more racially diverse make-up of voters have shifted the genres on the 2020 list compared to the 2003 one?

---

```{r, echo=FALSE}
genre_counts_2003 <- rolling_stone_2003 %>%
  group_by(genre) %>%
  summarise(album_count = n()) %>%
  arrange(desc(album_count))
```

```{r, echo=FALSE}
genre_counts_2020 <- rolling_stone_2020 %>%
  mutate(genre = str_to_title(str_trim(genre))) %>%
  group_by(genre) %>%
  summarise(album_count = n()) %>%
  arrange(desc(album_count))
```

```{r, echo=FALSE}
genre_counts_2003 <- rolling_stone_2003 %>%
  mutate(genre = str_to_title(str_trim(genre))) %>%
  group_by(genre) %>%
  summarise(album_count = n()) %>%
  arrange(desc(album_count))
```

```{r, echo=FALSE}
ggplot(genre_counts_2003, aes(x = genre, y = album_count, fill = genre)) +
  geom_bar(stat = "identity") +
  labs(title = "Genre Distribution of Artists (2003 List)",
       x = "Genre",
       y = "Album Count") +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = "none")

```

- New Wave and other Rock-adjacent genres appear most
- R&B appears quite frequently
- Reggae, Funk, Hip-hop, and Latin do not appear very often

---

```{r, echo=FALSE}
ggplot(genre_counts_2020, aes(x = genre, y = album_count, fill = genre)) +
  geom_bar(stat = "identity") +
  labs(title = "Genre Distribution of Artists (2020 List)",
       x = "Genre",
       y = "Album Count") +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = "none")

```

- Soul/Gospel/R&B now appears most often, above New Wave and other Rock genres
- Afrobeat added
- Hip-hop has over doubled and Funk, and Latin appear more often
---
class: inverse, middle, center

# Tables for Genre Make-up of Each List
---
.pull-left[
```{r, echo = FALSE}
kable(head(genre_counts_2003), format = "html")
```
]
.pull-right[
```{r, echo = FALSE}
kable(head(genre_counts_2020), format = "html")
```
]

.pull-left[
# 2003 list top 5 (plus N/A)
]

.pull-right[
# 2020 list top 5 (plus N/A)
]

---

.pull-left[
```{r, echo=FALSE}
genre_counts_2003 <- genre_counts_2003 %>%
  mutate(percentage = percent(album_count / sum(album_count)))
```

```{r, echo=FALSE}
genre_counts_2003 %>%
  head(6) %>%
  kable(format = "html", col.names = c("Genre", "Album Count", "Percentage"))
```
]

.pull-right[
```{r, echo=FALSE}
genre_counts_2020 <- genre_counts_2020 %>%
  mutate(percentage = percent(album_count / sum(album_count)))
```

```{r, echo=FALSE}
genre_counts_2020 %>%
  head(6) %>%
  kable(format = "html", col.names = c("Genre", "Album Count", "Percentage"))
```
]

.pull-left[
# 2003 list (with percentages)
]

.pull-right[
# 2020 list (with percentages)
]

---

class: inverse, middle, center

# Top 10 - 2003 vs 2020
---

# 2003 Top Ten

```{r, echo=FALSE}
rolling_stone_2003 %>%
  arrange(rank_2003) %>%
  select(rank_2003, album, clean_name, genre) %>%
  head(10) %>%
  kable(format = "html", col.names = c("Rank", "Album", "Artist", "Genre"))
```
---
# 2020 Top Ten

```{r, echo=FALSE}
rolling_stone_2020 %>%
  arrange(rank_2020) %>%
  select(rank_2020, album, clean_name, genre) %>%
  head(10) %>%
  kable(format = "html", col.names = c("Rank", "Album", "Artist", "Genre"))
```

---
class: inverse, middle, center

# Conclusion

---

# Based on the results, we can say that it appears that the gender and racial diversity of voters did affect the change in album placements between the 2003 and 2020 Rolling Stone lists.  There are more female artists and more genres connected to minority racial/ethnic groups in the 2020 list.

---

# However, this presentation is flawed/limited since there is no voter data in the `rolling_stone` dataset.  For example, we cannot see how many more female/minority voters voted in the 2020 list versus the 2003 list.  If we had this data, we could back up the results more effectively than by simply saying that there were more of these kinds of voters.

---

class: inverse, middle, center

# Sources

---

.pull-left[
https://pudding.cool/2024/03/greatest-music/
]

---

# Thank You

```{r music, echo = FALSE, out.width = "65%", fig.align = "center", fig.cap = "Image credit: Photo by blocks on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1483412033650-1015ddeb83d1?q=80&w=2946&auto=format&fit=crop&ixlib=rb-4.0.3&ixid=M3wxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8fA%3D%3D")
```
